---
title: "ES 193DS - homework 4"
format: html
author: Jasmine Toni
execute: 
  warning: false
  message: false
---

# ES 193DS Homework #4

```{r}
# should haves
library(tidyverse)
library(here)
library(lterdatasampler)

# would be nice to have
#install.packages("performance")
library(performance)
#install.packages("broom")
library(broom)
#install.packages("flextable")
library(flextable)
#install.packages("ggeffects")
library(ggeffects)
#install.packages("car")
library(car)
library(naniar)
```

```{r}
#load in the data from the website
fish <- read_csv(
  here("data", "ntl6_v12.csv")
)
```

Filter the data

```{r}
#filter the data we downloaded to look at the variables we want
trout_perch <- fish %>%
  filter(spname == "TROUTPERCH")
```

# Problem 1

1.  Mathematical Hypothesis:

    Null Hypothesis (H₀): There is no linear relationship between fish length and fish weight for trout perch across all sample years.

    Alternative Hypothesis (H₁): There is a linear relationship between fish length and fish weight for trout perch across all sample years.

    Biological Hypotheses:

    Null Hypothesis (H₀): Fish length does not predict fish weight for trout perch across all sample years.

    Alternative Hypothesis (H₁): Fish length predicts fish weight for trout perch across all sample years.

2.  Create a visual

    ```{r missing-data-vis}
    gg_miss_var(trout_perch)
    ```

a. The missing data is relevant to my hypothesis because there is some missing data in "weight", meaning that we will have some lengths without corresponding weights. This will creates problems because the data that I am observing will be incomplete and perhaps skew the results of the statistical tests and analysis.

3.  Run your test

```{r explore-data-vis}
# will show us our missing data visually (what categories everything is missing in)
ggplot(data = trout_perch, aes(x = length, y = weight)) +
  geom_point()
```

```{r linear-model-trout}
# making our linear model for the trout data
# will show us the intercept and length
trout_model <- lm(weight ~ length, data = trout_perch)

trout_model
```

4.  Visually check your assumptions

```{r checking-assumptions}
# checking the assumptions for the data
par(mfrow = c(2, 2))
plot(trout_model)
```

5.  Diagnostic Plot Analysis

Residuals vs Fitted Plot: This plot checks the linear relationship of the data and shows the red line (the mean of the dots and their line) fitting somewhat closely to the grey dotted line.

Normal Q-Q: This plot shows the points fitting on the grey dotted (expected line) very nicely, but tapering off a little bit at the end. This tells us that the data is relatively normally distributed.

Scale-Location: This plot tells us about the errors that should be constant between the models assumption of constant variance. In this case, it is homoskedasticity and indicates the variability is constant and the variability of the residuals is consistent across predicted values.

Residuals vs Leverage: This plot is a diagnostic plot used to assess the influence of individual data points on the regression model. Furthermore, Cook's distance demonstrates that a few outliers have skewed this plot and removing those specific observations could potentially have a substantial impact on the regression results.

6.  Summary results

```{r model-summary-table}
# store the model summary as an object
model_summary <- summary(trout_model)

# store the ANOVA table as an object
# anova(): special function to get analysis of variance tables for a model
model_squares <- anova(trout_model)

model_summary
```

7.  Creating a table

```{r}
# can't name this chunk!

model_squares_table <- tidy(model_squares) %>% 
  # round the sum of squares and mean squares columns to have 5 digits (could be less)
  mutate(across(sumsq:meansq, ~ round(.x, digits = 5))) %>% 
  # round the F-statistic to have 1 digit
  mutate(statistic = round(statistic, digits = 1)) %>% 
  # replace the very very very small p value with < 0.001
  mutate(p.value = case_when(
    p.value < 0.001 ~ "< 0.001"
  )) %>% 
  # rename the stem_length cell to be meaningful
  mutate(term = case_when(
    term == "Weight" ~ "Length",
    TRUE ~ term
  )) %>% 
  # make the data frame a flextable object
  flextable() %>% 
  # change the header labels to be meaningful
  set_header_labels(df = "Degrees of Freedom", 
                    sumsq = "Sum of squares",
                    meansq = "Mean squares",
                    statistic = "F-statistic",
                    p.value = "p-value")

model_squares_table
```
